A Decision Procedure for (Co)datatypes in SMT Solvers

### 2014-12-02 tex/codatadp.tex
more

  * initial algebra
    * ``no junk, no confusion''
    * initial algebra ``exists'' only if well-founded
      otherwise it would be empty, which is disallowed by most formulations of
      first-order logic (including CVC4 and other SMT solvers)
      and higher-order logic (including Isabelle/HOL)

@@

Bosses: Viktor Kuncak, Tobias Nipkow, and Stephan Merz.

>>>>

Axiomatization:

discriminators go here

specifies the

    * ``no junk, no confusion''

Because the types are inhabited (nonempty), a datatype specification is
admissible only if a ground term can be exhibited that belongs to that datatype.
This rules out ill-founded specifications such as
%
\[\keyw{datatype}\;\, \ty{fstream} \,=\, \const{FSCons}(\ty{int},\,\ty{fstream})\]
%
For codatatypes, no such check is necessary. Given the specification
%
\[\keyw{codatatype}\;\, \ty{stream} \,=\, \const{SCons}(\ty{int},\,\ty{stream})\]
%
the infinite value $\const{SCons}(0, \const{SCons}(0, \ldots))$ witnesses
nonemptiness.

Datatypes can be defined semantically as the initial model of an equational
specification of the selector--constructor equations. This is the view followed
by Barrett et al.\ \cite{barrett-et-al-2010}. A disadvantage with this approach
is that it does not naturally account for selectors applied to wrong
constructors. Barrett et al.\ address this by parameterizing the construction by
default values, but these give rise to spurious equalities between unrelated
terms (e.g., $hd(nil1) \teq hd(nil2)$). This could be corrected, but the added
compexity seems to suggest that selectors are better specified axiomatically.

A related semantic view of datatypes is as initial algebra. Codatatypes are then
defined dually as a final coalgebra. The datatypes are generated by their
constructors, whereas the codatatypes are viewed through their selectors
\cite{xxx}. By uniformly focusing on the constructors, the axiomatic approach
emphasizes the commonality between datatypes and codatatypes, while sacrificing
a theoretically fruitful notion of duality.

@@

Bosses: Viktor Kuncak, Stephan Merz, Tobias Nipkow, and Christoph Weidenbach.
